Safe and Robust Formal Synthesis of Digital Controllers for Continuous Plants

### 2018-03-07 automatica/paper.tex
Integrated Dario's changes into the paper
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

%% Using state of the art verifiers, we evaluate these models against
%% automatically synthesised controllers using CEGIS-based synsthesis that
%% incorporate the above mentioned artifacts in different models of
%% Single-Input/Single-Output (SISO) systems.
In this paper we present two novel approaches for the synthesis of digital controllers, 
both of which use a CounterExample-Guided Inductive Synthesis (CEGIS)~\cite{jha-icse10,
DBLP:conf/pldi/Solar-LezamaRBE05} as described next.

@@

The linear feedback controller is characterized by a gain matrix
$\mat{K}$ that influences the closed-loop dynamics
(Fig.~\ref{fig:digitalsystem}).  Later, we will address the issue of
variable quantization, as introduced by the ADC/DAC conversion blocks.

@@

We recall a technique called \emph{Abstract Acceleration}~\cite{JSS14,cattaruzza2015unbounded}, 
which will be key for abstraction-based synthesis.  
\note{This clashes with the model introduced above, with inputs - motivate why?}
Given a model of an iterative program with dynamics
\note{Pls eliminate unnecessary ``logical'' notation - we have discussed this already for the CAV18 submission, do we need to go through this again????}
%
\begin{equation}
\vec{x}_{k+1}=\mat{A}\vec{x}_k : \vec{x}_k \in \mathbb{R}^n \wedge \mat{A} \in \mathbb{R}^{n \times n} \wedge k \in \mathbb N_0,
\end{equation}
%
\note{material has been blindly pasted here - dimensions are all screwed up.}
we want to find all possible states visited over an infinite time horizon
starting at an initial set $X_0$.  
This set is known as the \emph{reach tube} of
the model:
%
\begin{equation}
\hat{X} = \{ \vec{x}_k : k \in \mathbb N_0 \wedge \vec{x}_0 \in X_0 \wedge \vec{x}_{k+1}=\mat{A}\vec{x}_k\}.
\end{equation}
%
It is easy to see that the infinite number of iterations would result in an
unbounded cumulative rounding error \note{why do we bring in the discussion on the rounding error here? why do we claim AA solves it? Is this the best manner to motivate the AA technique?}, but using Abstract Acceleration we
transform this problem into a one-step solution.  We first change the
problem statement by means of acceleration into
%
\begin{equation}
\hat{X} = \{ \vec{x}_k=\mat{A}^k\vec{x}_0 : k \in [0\ \infty) \wedge \vec{x}_0 \in X_0 \}, 
\end{equation}
%
and further define new model semantics and corresponding set: 
%
\begin{equation}\label{eq:aa_reachtube}
\hat{X} \subseteq \hat{X}^\sharp = \mathcal{A}X_0, \text{ such that } \bigcup_{k \in [0\ \infty)} \mat{A}^k \subseteq \mathcal{A}, 
\end{equation}
%
where the newly introduced \emph{abstract reach tube} $\hat{X}^\sharp$ is an
over-approximation of the reach tube of the model \cite{JSS14}.  All we need
to do is define the nature of $\mathcal{A}X_0$.

@@

Introduce set $X=\{ \vec{x} \mid \mat{C}\vec{x} \leq \vec{d}\}$. 
\note{Why are we interested in this set? What does this set represent?}
The linear transformation $\vec{x}'=\mat{S}^{-1}\vec{x}$ leads to set 
%
\begin{equation}
X'= \mat{S}^{-1}X = \{\vec{x}' : \mat{C}\mat{S}\vec{x}'\leq \vec{d}\}.  
\label{eq:eigenInit}
\end{equation}
%
Analogously, let $X_0'=\mat{S}^{-1}X_0$ and
${\hat{X}'^\sharp}=\mat{S}^{-1}{\hat{X}^\sharp}$ be the initial set and the
abstract reach tube mapped onto the eigenspace of matrix $\mat{A}$.  We may
transform equation \eqref{eq:aa_reachtube} into: 
%
\begin{equation}
\hat{X}' \subseteq {\hat{X}'^\sharp} = \mathcal{J}X_0', 
\end{equation}
%
where $\hat{X}'=\mat{S}^{-1}\hat{X}$ and $\mathcal{J}$ is a convex
polyhedron representing restrictions on the eigenvalues of $\mat{A}$. 
$\hat{X}'^\sharp$ is calculated by applying a simplex algorithm with a
tableau defined by $\mathcal{J}$ and objective functions derived from each
vertex of $X_0'$ (step \eqref{alg:getReachTube} in Algorithm
\ref{alg:findReachTube} below).  We note, however, that if we wish to use
numerical analysis in the decomposition of $\mat{A}$, the Tableau in
$\mathcal{J}$ contains small intervals given by the errors, which cannot be
processed efficiently by a regular simplex and require a new procedure.

The calculation of the abstract reach tube via abstract acceleration is
encompassed in Algorithm~\ref{alg:findReachTube}, which can be summarised as
follows (its steps are denoted in parentheses):
%
\begin{enumerate}
%
\item We perform unsound eigen-decomposition using an existing algebra package (lines \ref{alg:calcEigenvalues}-\ref{alg:calcEigenvectors}).
%
\item Then we restore soundness to the results using the methods described in \cite{cattaruzza2017sound} (lines \ref{alg:soundEigenvalues}-\ref{alg:soundEigenvectors}).
%
\item The inverse of the matrix of eigenvectors is calculated after soundness is restored, using interval arithmetic in order to ensure its soundness (line \ref{alg:inverse}).
%
\item The abstract dynamics are obtained by evaluating the convex hull of all powers of eigenvalues up to the desired number of iterations as described in \cite{cattaruzza2015unbounded} (line \ref{alg:aa}). 
%
\item Using Equation \eqref{eq:eigenInit}, transform the initial state into the eigenspace (line \ref{alg:eigeninit}).
%
\item Extract the vertices of the eigen-polyhedron $X_0'$ using the algorithms described in \cite{cattaruzza2015unbounded,cattaruzza2017sound} (line \ref{alg:getVertices}). 
%
\item Calculate the mapped \emph{abstract reach tube} ${\hat{X}'^\sharp}$
that over-approximates the image of the reach tube in the eigenspace (line
\ref{alg:getReachTube}).  This is achieved by evaluating a set of objective
functions using the abstract dynamics as a simplex Tableau, and via a sound
simplex described in \cite{cattaruzza2017sound}.  The objective functions are
defined by the vertices of $X'_0$, denote as $V_0$, and by the desired
template directions, such that
%
\begin{equation*}
\vec{w}_{ij}= \vec{v}_i \circ \vec{t}_j : \vec{v}_i \in V_0 \wedge \vec{t}_j \in T,
\end{equation*}
%
where $T$ is the set of template directions, and $\circ$ denotes a
component-wise multiplication yielding a vector.
%
\item Using Equation~\eqref{eq:eigenInit}, find the reach tube
$\hat{X}^\sharp=\mat{S}{\hat{X}'^\sharp}$ (line \ref{alg:getNormReachTube}).
%
\end{enumerate}

>>>>

%% Using state of the art verifiers, we evaluate these models against
%% automatically synthesised controllers using CEGIS-based synsthesis that
%% incorporate the above mentioned artifacts in different models of
%% Single-Input/Single-Output (SISO) systems.
In this paper we present two novel approaches for the synthesis of digital controllers, 
both of which use a CounterExample-Guided Inductive Synthesis (CEGIS)~\cite{%jha-icse10,
DBLP:conf/asplos/Solar-LezamaTBSS06} as described next.

@@

The linear feedback controller is characterized by a gain matrix
$\mat{K}$ that influences the closed-loop dynamics
(Fig.~\ref{fig:digitalsystem}) such that
\begin{equation*}
\vec{u}_k=\mat{K}\vec{x}_k \Rightarrow \vec{x}_{k+1}=(\mat{A}_d-\mat{B}_d\mat{K})\vec{x}_k
\end{equation*}
Which corresponds to a closed loop system without inputs. Later, we will address the issue of
variable quantization, as introduced by the ADC/DAC conversion blocks.

@@

We recall a technique called \emph{Abstract Acceleration}~\cite{JSS14,cattaruzza2015unbounded}, 
which will be key for abstraction-based synthesis.  
%\note{This clashes with the model introduced above, with inputs - motivate why? I've added the motivation above by introducing the closed  loop}
Given a model of an iterative program with dynamics
%
\begin{equation}
\vec{x}_{k+1}=\mat{A}\vec{x}_k \text{ where } \vec{x}_k \in \mathbb{R}^p, \mat{A} \in \mathbb{R}^{p \times p} \text{ and } k \in \mathbb N_0,
\end{equation}
%
we want to find all possible states visited over an infinite time horizon
starting at an initial set $X_0$.  
This set is known as the \emph{reach tube} of
the model:
%
\begin{equation}
\hat{X} = \{ \vec{x}_k \mid k \in \mathbb N_0, \vec{x}_0 \in X_0 \text{ and } \vec{x}_{k+1}=\mat{A}\vec{x}_k\}.
\end{equation}
%
Using Abstract Acceleration we
transform this problem into a one-step solution.  We first change the
problem statement by means of acceleration into
%
\begin{equation}
\hat{X} = \{ \vec{x}_k=\mat{A}^k\vec{x}_0 \mid k \in N_0 \text{ and } \vec{x}_0 \in X_0 \}, 
\end{equation}
%
and further define new model semantics and corresponding set: 
%
\begin{equation}\label{eq:aa_reachtube}
\hat{X} \subseteq \hat{X}^\sharp = \mathcal{A}X_0, \text{ such that } \bigcup_{k \in [0\ \infty)} \mat{A}^k \subseteq \mathcal{A}, 
\end{equation}
%
where the newly introduced \emph{abstract reach tube} $\hat{X}^\sharp$ is an
over-approximation of the reach tube of the model \cite{JSS14}.  All we need
to do is define the nature of $\mathcal{A}X_0$.

@@

Since abstract acceleration operates in the eigenspace, wemust deal with some preliminary transformations.
Given a set $X=\{ \vec{x} \mid \mat{C}\vec{x} \leq \vec{d}\}$,
%\note{Why are we interested in this set? What does this set represent?}
the linear transformation $\vec{x}'=\mat{S}^{-1}\vec{x}$ leads to the set 
%
\begin{equation}
X'= \mat{S}^{-1}X = \{\vec{x}' \mid \mat{C}\mat{S}\vec{x}'\leq \vec{d}\}.  
\label{eq:eigenInit}
\end{equation}
%
Analogously, let $X_0'=\mat{S}^{-1}X_0$ and
${\hat{X}'^\sharp}=\mat{S}^{-1}{\hat{X}^\sharp}$ be the initial set and the
abstract reach tube mapped onto the eigenspace of matrix $\mat{A}$.  We may
transform equation \eqref{eq:aa_reachtube} into: 
%
\begin{equation}
\hat{X}' \subseteq {\hat{X}'^\sharp} = \mathcal{J}X_0', 
\end{equation}
%
where $\hat{X}'=\mat{S}^{-1}\hat{X}$ and $\mathcal{J}$ is a convex
polyhedron representing restrictions on the eigenvalues of $\mat{A}$. 
$\hat{X}'^\sharp$ is calculated by applying a simplex algorithm with a
tableau defined by $\mathcal{J}$ and objective functions derived from each
vertex of $X_0'$ (step \eqref{alg:getReachTube} in Algorithm
\ref{alg:findReachTube} below).  We note, however, that if we wish to use
numerical analysis in the decomposition of $\mat{A}$, the Tableau in
$\mathcal{J}$ contains small intervals given by the errors, which cannot be
processed efficiently by a regular simplex and require a new procedure.

The calculation of the abstract reach tube via abstract acceleration is
encompassed in Algorithm~\ref{alg:findReachTube}, which can be summarised as
follows (its steps are denoted in parentheses):
%
\begin{enumerate}
%
\item We perform unsound eigen-decomposition using an existing algebra package (lines \ref{alg:calcEigenvalues}-\ref{alg:calcEigenvectors}).
%
\item Then we restore soundness to the results using the methods described in \cite{cattaruzza2017sound} (lines \ref{alg:soundEigenvalues}-\ref{alg:soundEigenvectors}).
%
\item The inverse of the matrix of eigenvectors is calculated after soundness is restored, using interval arithmetic in order to ensure its soundness (line \ref{alg:inverse}).
%
\item The abstract dynamics are obtained by evaluating the convex hull of all powers of eigenvalues up to the desired number of iterations as described in \cite{cattaruzza2015unbounded} (line \ref{alg:aa}). 
%
\item Using Equation \eqref{eq:eigenInit}, transform the initial state into the eigenspace (line \ref{alg:eigeninit}).
%
\item Extract the vertices of the eigen-polyhedron $X_0'$ using the algorithms described in \cite{cattaruzza2015unbounded,cattaruzza2017sound} (line \ref{alg:getVertices}). 
%
\item Calculate the mapped \emph{abstract reach tube} ${\hat{X}'^\sharp}$
that over-approximates the image of the reach tube in the eigenspace (line
\ref{alg:getReachTube}).  This is achieved by evaluating a set of objective
functions using the abstract dynamics as a simplex Tableau, and via a sound
simplex described in \cite{cattaruzza2017sound}.  The objective functions are
defined by the vertices of $X'_0$, denote as $V_0$, and by the desired
template directions, such that
%
\begin{equation*}
\vec{w}_{ij}= \vec{v}_i \circ \vec{t}_j \text{ with } \vec{v}_i \in V_0 \text{ and } \vec{t}_j \in T,
\end{equation*}
%
where $T$ is the set of template directions, and $\circ$ denotes a
component-wise multiplication yielding a vector.
%
\item Using Equation~\eqref{eq:eigenInit}, find the reach tube
$\hat{X}^\sharp=\mat{S}{\hat{X}'^\sharp}$ (line \ref{alg:getNormReachTube}).
%
\end{enumerate}
